Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.

### 2019-02-28 papers/CADE19/paper.tex
changes

The algorithm is centred around the \emph{clause selection} process. As previously mentioned, there are two main heuristics for this:
\begin{itemize}
	\item \emph{By Age (or First-in/First-out)} clause selection prefers the oldest clause, simulating a \emph{breadth-first} search of the clause space.
	\item \emph{By Weight (or symbol-counting)} clause selection prefers the youngest clause (with the fewest symbols). The intuition behind this approach is that the sought empty clause has zero symbols and lighter clauses are (in some sense) closer to this. Furthermore, lighter clauses are friendlier (less explosive) to proof search in various ways e.g. they are more general in terms of subsumption and tend to have fewer children.
\end{itemize}
It has been shown (see \sandm{}) that alternating these heuristics is beneficial and we consider this setting here. In Vampire this alternation is achieved by an age/weight ratio (AWR) implemented by a simple \emph{balancing} algorithm. The balance is initialised to 0 and used as follows: a positive balance means that a clause should be selected by age, whereas a negative balance means that a clause should be selected by weight; given a ratio of $a:w$ the balance is incremented by $a$ when selecting by age and decremented by $w$ when selecting by weight.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Optimising Age/Weight Ratios} \label{sec:awr:study}

@@

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but does follow an approximate trend and settles toward the global optimum.
These properties are typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

This experiment was repeated on the whole TPTP problem set, excluding problems Vampire does not support, problems which were not solved by the default mode in under a second for any AWR value, and problems solved in pre-processing.
The whole set yielded similar AWR data for 7,947 problems.
These data show that while there is no ``best'' AWR for the whole set (Figure \ref{fig:no-best-awr}), choosing a good AWR value for a problem is well-rewarded (Figure \ref{fig:optimal-awr-improvement}).

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{l l l l l}
AWR & Maximum & Mean & Standard Deviation\\
\hline
8:1 & 150.67 & 1.37 & 1.98\\
5:1 & 122.22 & 1.33 & 1.64\\
4:1 & 101.44 & 1.32 & 1.42\\
3:1 & 105.00 & 1.29 & 1.41\\
2:1 & 112.67 & 1.27 & 1.45\\
3:2 & 119.89 & 1.27 & 1.51\\
5:4 & 125.00 & 1.26 & 1.55\\
1 & 133.56 & 1.26 & 1.63\\
2:3 & 147.67 & 1.28 & 1.79\\
2 & 162.67 & 1.31 & 1.97\\
3 & 190.56 & 1.37 & 2.30\\
4 & 17.41 & 1.38 & 0.67\\
5 & 133.67 & 1.44 & 1.70\\
6 & 106.44 & 1.47 & 1.46\\
7 & 104.89 & 1.49 & 1.44\\
8 & 101.33 & 1.53 & 1.45\\
10 & 101.78 & 1.60 & 1.53\\
12 & 101.67 & 1.65 & 1.62\\
14 & 103.00 & 1.70 & 1.75\\
16 & 101.33 & 1.74 & 1.76\\
20 & 100.89 & 1.82 & 1.94\\
24 & 101.33 & 1.89 & 2.08\\
28 & 99.22 & 1.96 & 2.24\\
32 & 100.00 & 1.99 & 2.36\\
40 & 98.78 & 2.09 & 2.64\\
50 & 98.78 & 2.17 & 2.88\\
64 & 101.22 & 2.28 & 3.19\\
128 & 107.44 & 2.57 & 4.16\\
1024 & 222.01 & 2.83 & 7.55
	\end{tabular}
	\caption{The mean relative performance drop (in terms of activations compared to an optimal value) choosing a single AWR over an entire problem set causes. No setting is significantly ``better'' or ``worse'' than the others, but some have more variance in performance. A ratio of 1:4 produces an unusually well-behaved maximum performance drop. \sandm~found that 1:5 had a similar property, but this might be explained by differences in test environment.}
	\label{fig:no-best-awr}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variable AWR for Vampire}

>>>>

The algorithm is centred around the \emph{clause selection} process. As previously mentioned, there are two main heuristics for this:
\begin{itemize}
	\item \emph{By Age (or First-in/First-out)} clause selection prefers the oldest clause (produced earlier in proof search), simulating a \emph{breadth-first} search of the clause space. In Vampire the age of a clause is the number of inferences performed to produce it (e.g. input clauses have age 0).
	\item \emph{By Weight (or symbol-counting)} clause selection prefers the youngest clause (with the fewest symbols). The intuition behind this approach is that the sought empty clause has zero symbols and lighter clauses are (in some sense) closer to this. Furthermore, lighter clauses are friendlier (less explosive) to proof search in various ways e.g. they are more general in terms of subsumption and tend to have fewer children.
\end{itemize}
It has been shown (see \sandm{}) that alternating these heuristics is beneficial and we consider this setting here. In Vampire this alternation is achieved by an age/weight ratio (AWR) implemented by a simple \emph{balancing} algorithm. The balance is initialised to 0 and used as follows: a positive balance means that a clause should be selected by age, whereas a negative balance means that a clause should be selected by weight; given a ratio of $a:w$ the balance is incremented by $a$ when selecting by age and decremented by $w$ when selecting by weight.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage~\newpage
\section{Optimising Age/Weight Ratios} \label{sec:awr:study}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but follows an approximate trend and settles toward the global optimum.
This is typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

@@

This experiment was repeated on the whole TPTP problem set, excluding problems Vampire does not support, problems which were not solved by the default mode in under a second for any AWR value, and problems solved in pre-processing.
The whole set yielded similar AWR data for 7,947 problems.
These data show that while there is no ``best'' AWR for the whole set (Table \ref{tab:no-best-awr}), choosing a good AWR value for a problem is well-rewarded (Figure \ref{fig:optimal-awr-improvement}).

@@

\begin{table}[t]
	\caption{The mean relative performance drop (in terms of activations compared to an optimal value) choosing a single AWR over an entire problem set causes. No setting is significantly ``better'' or ``worse'' than the others, but some have more variance in performance. A ratio of 1:4 produces an unusually well-behaved maximum performance drop. \sandm~found that 1:5 had a similar property, but this might be explained by differences in test environment.}
	\centering
	\begin{tabular}{l | c | c  c }
AWR & Maximum Drop & Mean Drop & (Standard Deviation)\\
\hline
8:1 & 150.67 & 1.37 & 1.98\\
5:1 & 122.22 & 1.33 & 1.64\\
4:1 & 101.44 & 1.32 & 1.42\\
3:1 & 105.00 & 1.29 & 1.41\\
2:1 & 112.67 & 1.27 & 1.45\\
3:2 & 119.89 & 1.27 & 1.51\\
5:4 & 125.00 & 1.26 & 1.55\\
1 & 133.56 & 1.26 & 1.63\\
2:3 & 147.67 & 1.28 & 1.79\\
2 & 162.67 & 1.31 & 1.97\\
3 & 190.56 & 1.37 & 2.30\\
4 & 17.41 & 1.38 & 0.67\\
5 & 133.67 & 1.44 & 1.70\\
6 & 106.44 & 1.47 & 1.46\\
7 & 104.89 & 1.49 & 1.44\\
8 & 101.33 & 1.53 & 1.45\\
10 & 101.78 & 1.60 & 1.53\\
12 & 101.67 & 1.65 & 1.62\\
14 & 103.00 & 1.70 & 1.75\\
16 & 101.33 & 1.74 & 1.76\\
20 & 100.89 & 1.82 & 1.94\\
24 & 101.33 & 1.89 & 2.08\\
28 & 99.22 & 1.96 & 2.24\\
32 & 100.00 & 1.99 & 2.36\\
40 & 98.78 & 2.09 & 2.64\\
50 & 98.78 & 2.17 & 2.88\\
64 & 101.22 & 2.28 & 3.19\\
128 & 107.44 & 2.57 & 4.16\\
1024 & 222.01 & 2.83 & 7.55
	\end{tabular}

	\label{tab:no-best-awr}
\end{table}

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variable AWR for Vampire}
